Meta-analytic approaches and 'new' effect sizes to account for heterogeneous effects in comparative physiology

### 2021-06-04 MS/ms.Rmd
Started writing heterogenity section some more

---
title: "Meta-analytic approaches and 'new' effect sizes to account for heterogeneous effects in comparative physiology"
bibliography: ../bib/refs.bib
csl: ../bib/the-journal-of-experimental-biology.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
  ## numbers >= 10^5 will be denoted in scientific notation,		  ## numbers >= 10^5 will be denoted in scientific notation,
    ## and rounded to 2 digits		  ## and rounded to 2 digits
    options(digits = 2)
```

@@

```{r tablea, echo = FALSE, tab.cap = "Common effect sizes, associated sampling variances and notes on when they might be used. Note that for log odds ratio a continuity correction is applied (i.e., 0.5) to each proportion. In addition, a correction factor, J (1-3 / ..), is applied to the standardised to correct for small sample effects"}

# I'm just listing a bunch here, but we should think about what we want to include. Here's an example of how we can make a table with all equations. Fun!
names <- c("Mean",
           "Variance",
           "Log Response Ratio",
           "Standardised Mean Difference",
           "Zr (Correlation Coefficient)",
           "Log Odds Ratio")
eqns_yi <- c("X_{i}",
             "\\sigma^2",
             "log \\left( \\frac{X_{1}}{X_{2}} \\right)",
             "\\frac{\\left( X_{1} - X_{2}  \\right)}{sd_{p}}J",
             "\\frac{1}{2} log \\left( \\frac{1+r_{i}}{1-r_{i}} \\right)",
             "log \\left( \\frac{y_{i2}+0.5}{n_{i2} - y_{i2}+0.5} \\right) - log \\left( \\frac{y_{i1}+0.5}{n_{i1} - y_{i1}+0.5} \\right)")
#eqns_vi <- c()

table_fin <- data.frame(names, eqns_yi)

# Note: It is esssential that you be clear that the function 'compose' is exported from the flextable package. That's because purr also has the same function, so it gets messy.
tablea <- flextable(table_fin) %>% 
          flextable::compose(j = "names", part = "header", value = as_paragraph(as_b("Effect Measure"))) %>% 
          flextable::compose(j = "eqns_yi", part = "header", value = as_paragraph(as_b("Definition"))) %>% 
          flextable::compose(j = "eqns_yi", value = as_paragraph(as_equation(eqns_yi))) %>% 
          flextable::width(j = 1, 3) %>% 
          flextable::width(j = 2, 2)

@@

#### *Heterogenity Analysis: How Much Effect Size Variation Actually Exists?*
Quantifying effect size heterogenity (i.e., effect size variance) provides a pathway to explore hypotheses thought to explain why effect might vary across studies.

>>>>

---
title: "Meta-analytic approaches and 'new' effect sizes to account for heterogeneous effects in comparative physiology"
bibliography: ../bib/refs.bib
csl: ../bib/the-journal-of-experimental-biology.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: no
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
  ## numbers >= 10^5 will be denoted in scientific notation,		  ## numbers >= 10^5 will be denoted in scientific notation,
    ## and rounded to 2 digits		  ## and rounded to 2 digits
    options(digits = 2)

  pacman::p_load(flextable, tidyverse, equatags)  
```

@@

```{r tablea, echo = FALSE, tab.cap = "Common effect sizes, associated sampling variances and notes on when they might be used. Note that for log odds ratio a continuity correction is applied (i.e., 0.5) to each proportion. In addition, a correction factor, J (1-3 ), is applied to the standardised to correct for small sample effects"}

# I'm just listing a bunch here, but we should think about what we want to include. Here's an example of how we can make a table with all equations. Fun! Really useful website here: https://math.meta.stackexchange.com/questions/5020/mathjax-basic-tutorial-and-quick-reference. Something wrong with character coding when rendering word document. Has to do with the log odds ratio equation. Renders fine in html, but just stops when word
names <- c("Mean",
           "Variance",
           "Log Response Ratio",
           "Standardised Mean Difference",
           "Zr (Correlation Coefficient)")
           #"Log Odds Ratio",
           #"")

eqns_yi <- c("X_{i}",
             "\\sigma^2",
             "log \\left( \\frac{X_{1}}{X_{2}}\\right)",
             "\\frac{\\left( X_{2} - X_{1}\\right)}{sd_{p}}J",
             "\\frac{1}{2}log\\left(\\frac{1+r_{i}}{1-r_{i}}\\right)")
             #"log\\left(\\frac{y_{i2}+0.5}{n_{i2}-y_{i2}+0.5}\\right) ",
             #"- log\\left(\\frac{y_{i1}+0.5}{n_{i1}-y_{i1}+0.5}\\right)")

# Just placeholders for now
eqns_vi <- c("\\frac{s_{i}^2}{n_{i}}",
             "\\frac{s_{i}^2}{n_{i}}",
             "\\frac{s_{i}^2}{n_{i}}",
             "\\frac{s_{i}^2}{n_{i}}",
             "\\frac{s_{i}^2}{n_{i}}")
             #"\\frac{s_{i}^2}{n_{i}}",
             #"")

table_fin <- data.frame(names, eqns_yi, eqns_vi)

# Note: It is esssential that you be clear that the function 'compose' is exported from the flextable package. That's because purr also has the same function, so it gets messy.
tablea <- flextable(table_fin)                                                                                %>% 
          flextable::compose(j = "names", part = "header", value = as_paragraph(as_b("Effect Measure")))      %>% 
          flextable::compose(j = "eqns_yi", part = "header", value = as_paragraph(as_b("Definition")))        %>% 
          flextable::compose(j = "eqns_vi", part = "header", value = as_paragraph(as_b("Sampling Variance"))) %>%
          flextable::compose(j = "eqns_yi", value = as_paragraph(as_equation(eqns_yi)))                       %>%
          flextable::compose(j = "eqns_vi", value = as_paragraph(as_equation(eqns_vi)))                       %>% 
          flextable::width(j = 1, 3)                                                                          %>% 
          flextable::width(j = 2, 2)

@@

#### *Heterogenity Analysis: How Much Effect Size Variation Actually Exists?*
Quantifying effect size heterogenity (i.e., effect size variance) provides a pathway to explore hypotheses thought to explain why effects might vary within and across studies. Effect size heterogeneity can be thought of as the percentage of variance between effect sizes that cannot be attributable to sampling error (often as called 'total' heterogenity) @Higgins2002. It provides a formal way to assess how much variation is the result of 'true' biological or methodological differences as opposed to simply chance (i.e., sampling variance). We can only conduct such analyses if we have an estimate of the sampling variance for a given effect measure becasue only then can we quantify how much variation is the result of sampling variability. @Senior2016 has shown that total heterogenity is often extremely high in ecological and evolutionary meta-analysis (~91%), suggesting that effect measures vary widely across within and studies. Some of the variation is clearly the result of working with different species or strains, different traits and different outcome measures, however, some of this variation also arises because of methodological differences across studies
